Application of Generalized Additive Models for long-term and seasonal trend analysis of water quality: South San Francisco Bay case study

### 2020-06-10 manu_draft.Rmd
trends map done

data(locs)
data(modprf)
data(seastrnd)
data(mods_chl18)
data(mods_chl24)
data(mods_chl27)
data(mods_chl32)
data(mods_chl34)

@@

Table \@ref(tab:modprftab), Table \@ref(tab:seastrndtab), Figure \@ref(fig:prddat), Figure \@ref(fig:trnddat)

@@

>>>>

data(locs)
data(modprf)
data(seastrnd)
data(seastrnd2)
data(chgtrnd)
data(mods_chl18)
data(mods_chl24)
data(mods_chl27)
data(mods_chl32)
data(mods_chl34)

@@

Table \@ref(tab:modprftab), Table \@ref(tab:seastrndtab), Figure \@ref(fig:prddat), Figure \@ref(fig:trnddat), Figure \@ref(fig:trndmap)

@@

```{r, results = 'hide', message = F}
# change opacity of basemap
mapatt <- attributes(bsmap1)
bsmap1_transparent <- matrix(adjustcolor(bsmap1, 
                                         alpha.f = 0.4), 
                             nrow = nrow(bsmap1))
attributes(bsmap1_transparent) <- mapatt

leglab2 <- expression(paste("Chl-a change (", italic(mu), "g ", L^-1, yr^-1, ")"))

pthm <- theme_bw(base_family = 'serif', base_size = 14) +
  theme(
    legend.position = 'top',
    strip.background = element_blank(),
    axis.title = element_blank(), 
    axis.text = element_text(size = 7)
  )

toplo1 <- chgtrnd %>% 
  mutate(
    station = as.numeric(station),
    pval = ifelse(pval == 'sig', '*', '')
    ) %>% 
  left_join(locs, by = 'station') %>% 
  filter(model %in% 'gam6')

toplo2 <- seastrnd2 %>%
  mutate(
    station = as.numeric(station),
    pval = ifelse(pval < 0.05, '*', ''), 
    coefsgn = sign(yrcoef), 
    coefsgn = factor(coefsgn, levels = c('1', '-1'), labels = c('inc', 'dec'))
  ) %>% 
  left_join(locs, by = 'station')

p1 <- ggmap(bsmap1_transparent) +
  geom_point(data = toplo1, aes(x = lon, y = lat, size = abs(perchg), shape = persgn, fill = perchg, colour = pval)) + #, color = 'black') +
  geom_text(data = toplo1, aes(x = lon, y = lat, label = pval), nudge_x = 0.04) + 
  facet_grid( ~ yrs) + 
  scale_fill_gradient2('Chl-a % change', low = 'green', mid = 'white',  high = 'tomato1', midpoint = 0) +
  scale_colour_manual(values = c('black', 'black'), guide = F) +
  coord_map() + 
  scale_shape_manual('Trend', values = c(24, 25)) + 
  pthm +
  scale_size(range = c(1, 6), guide = F) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 6)) +
  labs(
    subtitle = '(a) Percent change between year pairs'
  )

p2 <- ggmap(bsmap1_transparent) +
  geom_point(data = toplo2, aes(x = lon, y = lat, size = abs(yrcoef), shape = coefsgn, fill = yrcoef, colour = pval)) + #, color = 'black') +
  geom_text(data = toplo2, aes(x = lon, y = lat, label = pval), nudge_x = 0.04) + 
  facet_grid(seas ~ yrs) + 
  scale_fill_gradient2(leglab2, low = 'green', mid = 'white',  high = 'tomato1', midpoint = 0) +
  scale_colour_manual(values = c('black', 'black'), guide = F) +
  coord_map() + 
  scale_shape_manual('Trend', values = c(24, 25)) + 
  pthm +
  scale_size(range = c(1, 6), guide = F) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 6)) +
  labs(
    subtitle = '(b) Annual rates of change across seasons by decade'
  )

p <- p1 + p2 + plot_layout(ncol = 1, heights = c(0.5, 1))

png('figs/trndmap.png', height = 10, width = 6.5, family = 'serif', units = 'in', res = 400)
p
dev.off()
```
```{r trndmap, fig.cap= "Trend estimates (a) between year pairs and (b) across seasons by decade for chlorophyll at each station.  The top plots show percent change comparing the first and last year within a decade and the bottom plots show seasonal estimates of change per year for chlorophyll concentrations for each decade.  Point type, shape, and color represent the direction and magnitude of an estimated trend.  Trends with $p<0.05$ are marked with an asterisk.  All results are from `gam6`."}
knitr::include_graphics('figs/trndmap.png')
```
